### 2017-09-05 analyses/Loading_data_example.Rmd
Initial commit

---
title: "quick pipeline for getting data from discrete matrices"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

Quick pipeline for getting data from discrete matrices.
===============

Packages
-----------------------------------

Make sure you work with `dispRity` 0.4 and my version of `Claddis` (that is much faster than Graeme's).

```{r, eval = TRUE}
# The latest version of devtools and ape
if(!require(devtools)) install.packages("devtools")
if(!require(ape)) install.packages("ape")

## The latest version of Claddis (on development on my own branch)
install_github("TGuillerme/Claddis")

## The latest version of dispRity 
install_github("TGuillerme/dispRity", ref = "release")
```

Loading the data
-----------------------------------

Loading the packages and setting the seed:

```{r}
set.seed(123)
library(ape); library(Claddis) ; library(dispRity) 
```

Loading the data.
It's based on a chain name (here "Beck2014") and should work with different chain names if their files are in the exact same format (it's not the case yet!).

```{r}
## Setting up the chain name
chain <- "Beck2014"

## Loading the discrete morphological matrix
matrix <- ReadMorphNexus(paste0("../Data/Matrices/", chain, ".nex"))

## Loading the FADLAD
FADLAD <- read.csv(paste0("../Data/FADLAD/", chain, ".csv"), row.names = 1, header = TRUE)

## Loading the tree
tree <- read.nexus(paste0("../Data/Trees/", chain, ".tre"))

## Adding node labels
tree <- makeNodeLabel(tree, method = "number", prefix = "n")

## Adding a root age (max tree age date)
tree$root.time <- max(tree.age(tree)[,1])
```

We'll also clean the data to match the tree and the matrix:

```{r}
## Cleaning the data 
cleaned_data <- clean.data(matrix$matrix, tree)

## Cleaned matrix
matrix$matrix <- cleaned_data$data

## Cleaned tree
tree <- cleaned_data$tree

## Removed taxa:
as.vector(na.omit(c(cleaned_data$dropped_tips, cleaned_data$dropped_rows)))
```

Ordination
===============

Ancestral states reconstruction
-----------------------------------

First let's recreate the nodal states but excluding any estimation with less than 0.95 scaled likelihood value.
This allows us to discard any dodgy/ambiguous reconstruction (note that NAs are not a problem since we use Gower's distances further down the road).

This have some running time!

```{r, eval = TRUE}
## Estimating the ancestral states (with a uncertainty threshold of 0.95)
matrix_nodes <- AncStateEstMatrix(matrix, tree, estimate.allchars = TRUE,
                                  uncertainty.threshold = 0.95)

## Saving the matrix
save(matrix_nodes, file = paste0("../data/backups/ancestral_states.", chain ,".Rda"))
```

This takes some time (couple of hours) so it can be skipped for the demo.

After that we can add the node states to the overall matrix:

```{r}
## Loading the matrix
load(paste0("../data/backups/ancestral_states.", chain ,".Rda"))

## Adding the node names as row names
row.names(matrix_nodes) <- tree$node.label

## Combining the tips and the nodes in the same matrix
matrix$matrix <- rbind(matrix$matrix, matrix_nodes) 
```

Distance matrix
-----------------------------------

Let's then calculate the distance matrix (using Gower's distance).
Same thing, the step takes some time (couple of minutes) so can be skipped for convenience.

```{r, eval = TRUE}
## Calculating the Gower distance
matrix_dist <- MorphDistMatrixFast(matrix, distance = "Gower")

## Saving the matrix
save(matrix_dist, file = paste0("../data/backups/distance_matrix.", chain ,".Rda"))
```

Ordination
-----------------------------------

Ordination is a classic NMDS including only the non-zero variance axis.

```{r}
## Loading the distance matrices
load(paste0("../data/backups/distance_matrix.", chain ,".Rda"))

## Ordinating the matrix
morphospace <- cmdscale(matrix_dist, k = nrow(matrix_dist) - 2, add = T)$points
```

And that's it... You know have all the arguments for running the `dispRity` package properly: the space (`morphospace`), the tree (`tree`) and the FADLAD (`FADLAD`).
Remember that for the `time.subsamples` function, you can use the `inc.nodes` logical option to include nodes or not!