### 2017-11-07 analyses/02-time-slice-analyses.Rmd
pipelined analysis (test on test data not real!) and plots

---
title: "Time slice analyses"
author: "Natalie Cooper and Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

This code runs the analyses in the paper. Below we lay out the process in detail for one dataset, and then provide the code for the other datasets below this. Note that much of this uses functions in the `simulations.R` file.

```{r}
set.seed(123)
```

Load packages and functions
-----------------------------------

Make sure you work with `dispRity` 0.4 and TG's version of `Claddis` (much faster than GL's).

```{r, eval = TRUE, message = FALSE}
# The latest version of devtools and ape
library(devtools)
library(ape)

## The latest version of Claddis (on development on TG branch)
## install_github("TGuillerme/Claddis")
library(Claddis)

## The latest version of dispRity on master branch
## install_github("TGuillerme/dispRity", ref = "master")
library(dispRity)

## Load functions using source
# Doesn't always work... Rmd problem?
source("../functions/simulations_fun.R")
```

Loading the data
-----------------------------------

To run `dispRity` analyses we need the **morphospace**, the **phylogeny** (with a root age and node labels) and the **first and last occurrence dates** (FADLAD). The tree and FADLAD are in the `data/` folder. To construct the morphospace we need to use the `distance_matrix` files we created in `01-extract-data-for analyses.Rmd` and stored in `data/processed`.

```{r}
## Choose a slug
slug <- "Beck2014"

## Loading the discrete morphological matrix
## This is just to match the names with the tree
matrix <- ReadMorphNexus(paste0("../data/matrices/", slug, ".nex"))

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Loading the tree
tree <- read.nexus(paste0("../data/trees/", slug, ".tre"))
## Adding node labels and a root age (max tree age date)
tree <- makeNodeLabel(tree, method = "number", prefix = "n")
tree$root.time <- max(tree.age(tree)[, 1])

## Cleaning the tree to remove things that aren't in the matrix 
cleaned_data <- clean.data(matrix$matrix, tree)
matrix$matrix <- cleaned_data$matrix
tree <- cleaned_data$tree

## Loading the distance matrices
## See "01-extract-data-for analyses.Rmd" for how we made these
load(paste0("../data/processed/distance_matrix.", slug ,".Rda"))

## Ordinating the matrix
morphospace <- cmdscale(matrix_dist, k = nrow(matrix_dist) - 2, add = TRUE)$points

```

Scenarios
--------------------------------

We used the data to explore these different scenarios:

1. Stratigraphy. This is the traditional method, where all the species within each stratigraphic period are included in the disparity calculation. This often leads to bins of unequal duration.

2. Equally sized time bins. 
    i. Where the *duration* of the bin is equal to the median duration of the stratigraphic period.
    ii. Where the *number* of bins is equal to the number of stratigraphic periods.
    iii. Where the *number* of bins is either 2, 5, 10, 15, or 20.

3. Time slicing. 
    i. Where the *interval* between slices is equal to the median duration of the stratigraphic period.
    ii. Where the *number* of slices is equal to the number of stratigraphic periods.
    iii. Where the *number* of slices is either 2, 5, 10, 15, or 20.

Running the analyses and saving the outputs
--------------------------------------------

These analyses use a series of functions in `functions\`. These essentially wrap the existing `dispRity` functions so that running lots of models and extracting lots of data is not such a big issue.

> TG: for the following pipeline, I've reduced the number of rarefaction that highly increased the calculation time.
We now only calculate the disparity for the non-rarefied subsamples and the subsamples rarefied to contain only 3 taxa (the lowest level of rarefaction).
We can use the full rarefaction if we want to plot rarefaction curves later on.

```{r}
## Testing data: TO REMOVE
#TG: it's just faster this way
#TG: also, I've only tested it on "Epoch" since it take ages to run, should be pipelined all right though
data(BeckLee_mat99)
data(BeckLee_tree)
data(BeckLee_ages)

morphospace <- BeckLee_mat99
tree <- BeckLee_tree
FADLAD <- BeckLee_ages

```

```{r, eval = FALSE, warning = FALSE}
## Run the function for Beck data
slug <- "BeckLee"

## Ages (rarefied to 3 taxa only)
out1 <- run.all.disparity.wrapper(morphospace, tree, type = "Age", FADLAD, inc.nodes = TRUE,
                                bootstraps = 100, metric = c(sum, variances),
                                metric.name = "sum_var", rarefaction = 3)
#readr::write_csv(out1, path = paste0("../outputs/age", "_", slug, ".csv"))

## Epochs (rarefied to 3 taxa only)
out2 <- run.all.disparity.wrapper(morphospace, tree, type = "Epoch", FADLAD, inc.nodes = TRUE,
                        bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var",
                        rarefaction = FALSE)
#readr::write_csv(out2, path = paste0("../outputs/epoch", "_", slug, ".csv"))
```

We also ran the models using varying numbers of bins as follows.

```{r, eval = FALSE}
## Run the function for Beck data
slug <- "BeckLee"

for (bins in c(2, 5, 10, 15, 20)){
  output <- run.all.disparity(morphospace, tree, bins, FADLAD, inc.nodes = TRUE,
                              bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var")

  readr::write_csv(output, path = paste0("../outputs/bins", "_", slug, ".csv"), append = TRUE)
}

```
<!-- Need to think more about the outputs here, currently they don't make much sense as there are no column headers -->

<!-- TG: Suggestion for a "testing the effect of binning/slicing" pipeline -->

Examining the output 
--------------------

The results as a table

```{r, eval = FALSE}
knitr::kable(out2$results$stratigraphy, caption = "Disparity per stratigraphic time: unequal time bins and non-equidistant time slices")
knitr::kable(out2$results$duration, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the median stratigraphic epoch duration")
knitr::kable(out2$results$number, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the number of stratigraphic epochs")
```

This is not really digest and we should find a better way to present that.
Anyway, these boring tables are always good for the supplementaries.

Plotting wise, we have to solutions, either using `dispRity`

```{r, eval = FALSE}
op <- par(bty = "n")
colors <- palette() #TG: that can be made fancier

## STRATIGRAPHY
## Plotting the continuous data
plot(out2$object$stratigraphy[[2]], quantile = 1, col = colors[1], ylim = c(1,4), main = "Equal duration and stratigraphys equal to stratigraphy", xlab = "sum of ranges")
plot(out2$object$stratigraphy[[3]], quantile = 1, col = colors[2], ylim = c(1,4), add = TRUE)
plot(out2$object$stratigraphy[[4]], quantile = 1, col = colors[3], ylim = c(1,4), add = TRUE)
plot(out2$object$stratigraphy[[5]], quantile = 1, col = colors[4], ylim = c(1,4), add = TRUE)
## Plotting the discrete data
plot(out2$object$stratigraphy[[1]], quantile = 1, type = "c", col = colors[5], ylim = c(1,4), add = TRUE, density = 0)
#TG: density = 0 is to not show the huge confidence interval, even for quantile = 1!
legend(x = "bottomright", legend = c("acctran", "deltran", "punctuated", "gradual", "time bins"), col = colors[1:5], lty = 1)
```

This on is interesting with no time bins after around 39 Mya and a drastic drop!

```{r, eval = FALSE}
op <- par(bty = "n")
colors <- palette() #TG: that can be made fancier
## DURATION
## Plotting the continuous data
plot(out2$object$duration[[2]], quantile = 1, col = colors[1], ylim = c(1,4), main = "Equal duration and durations equal to stratigraphy", xlab = "sum of ranges")
plot(out2$object$duration[[3]], quantile = 1, col = colors[2], ylim = c(1,4), add = TRUE)
plot(out2$object$duration[[4]], quantile = 1, col = colors[3], ylim = c(1,4), add = TRUE)
plot(out2$object$duration[[5]], quantile = 1, col = colors[4], ylim = c(1,4), add = TRUE)
## Plotting the discrete data
plot(out2$object$duration[[1]], quantile = 1, type = "c", col = colors[5], ylim = c(1,4), add = TRUE, density = 0)
legend(x = "bottomright", legend = c("acctran", "deltran", "punctuated", "gradual", "time bins"), col = colors[1:5], lty = 1)
```

Same absence of data after 39 but no drop!

```{r, eval = FALSE}
op <- par(bty = "n")
colors <- palette() #TG: that can be made fancier
## NUMBERS
## Plotting the continuous data
plot(out2$object$number[[2]], quantile = 1, col = colors[1], ylim = c(1,4), main = "Equal duration and numbers equal to stratigraphy", xlab = "sum of ranges")
plot(out2$object$number[[3]], quantile = 1, col = colors[2], ylim = c(1,4), add = TRUE)
plot(out2$object$number[[4]], quantile = 1, col = colors[3], ylim = c(1,4), add = TRUE)
plot(out2$object$number[[5]], quantile = 1, col = colors[4], ylim = c(1,4), add = TRUE)
## Plotting the discrete data
plot(out2$object$number[[1]], quantile = 1, type = "c", col = colors[5], ylim = c(1,4), add = TRUE)
legend(x = "bottomright", legend = c("acctran", "deltran", "punctuated", "gradual", "time bins"), col = colors[1:5], lty = 1)

```

This one is fairly similar but starts earlier in time.

Another way would be to do histograms the same way we drew it when we designed the project but I'll work on that tomorrow I guess.