### 2017-11-27 analyses/02-time-slice-analyses.Rmd
Remove examining output section and save otuputs instead

---
title: "Time slice analyses"
author: "Natalie Cooper and Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

This code runs the analyses in the paper. Below we lay out the process in detail for one dataset, and then provide the code for the other datasets below this. Note that much of this uses functions in the `simulations.R` file.

```{r}
set.seed(123)
```

Load packages and functions
-----------------------------------

Make sure you work with `dispRity` 0.4 and TG's version of `Claddis` (much faster than GL's).

```{r, eval = TRUE, message = FALSE}
# The latest version of devtools and ape
library(devtools)
library(ape)

## The latest version of Claddis (on development on TG branch)
## install_github("TGuillerme/Claddis")
library(Claddis)

## The latest version of dispRity on release branch #TG: master is under many changes!
## install_github("TGuillerme/dispRity", ref = "release")
library(dispRity)

## Load functions using source
source("../functions/simulations_fun.R")
```

Loading the data
-----------------------------------

To run `dispRity` analyses we need the **morphospace**, the **phylogeny** (with a root age and node labels) and the **first and last occurrence dates** (FADLAD). The FADLAD is in the `data/` folder, the tree with node labels and cleaned of taxa not in the morphospace is in the `data/processed` folder. To construct the morphospace we need to use the `distance_matrix` files we created in `01-extract-data-for analyses.Rmd` and stored in `data/processed`.

```{r}
## Choose a slug
slug <- "Beck2014"

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Loading the tree
## See "01-extract-data-for analyses.Rmd" for how we cleaned these
tree <- read.nexus(paste0("../data/processed/tree.", slug, ".tre"))
## Add root age (max tree age date)
tree$root.time <- max(tree.age(tree)[, 1])
```

Ordinating the matrix
----------------

Ordination is a classic NMDS (multidimensional scaling) including only the non-zero variance axes.

```{r}
## Loading the distance matrices
## See "01-extract-data-for analyses.Rmd" for how we made these
load(paste0("../data/processed/distance_matrix.", slug ,".Rda"))

## Ordinating the matrix
morphospace <- cmdscale(matrix_dist, k = nrow(matrix_dist) - 2, add = TRUE)$points
```

Scenarios
--------------------------------

We used the data to explore these different scenarios:

1. Stratigraphy. This is the traditional method, where all the species within each stratigraphic period are included in the disparity calculation. This often leads to bins of unequal duration.

2. Equally sized time bins. 
    i. Where the *duration* of the bin is equal to the median duration of the stratigraphic period.
    ii. Where the *number* of bins is equal to the number of stratigraphic periods.
    iii. Where the *number* of bins is either 2, 5, 10, 15, or 20.

3. Time slicing. 
    i. Where the *interval* between slices is equal to the median duration of the stratigraphic period.
    ii. Where the *number* of slices is equal to the number of stratigraphic periods.
    iii. Where the *number* of slices is either 2, 5, 10, 15, or 20.

Running the analyses and saving the outputs
--------------------------------------------

These analyses use a series of functions in `functions\`. These essentially wrap the existing `dispRity` functions so that running lots of models and extracting lots of data is not such a big issue. The function `run.all.disparity.wrapper` will run all the scenarios above.

```{r, eval = TRUE, warning = FALSE}
## Run the function for Beck data
slug <- "Beck2014"

## Ages (rarefied to 3 taxa only)
out1 <- run.all.disparity.wrapper(morphospace, tree, type = "Age", FADLAD, inc.nodes = TRUE,
                                bootstraps = 100, metric = c(sum, variances),
                                metric.name = "sum_var", rarefaction = FALSE)

## Save results to output folder
save(out1, file = paste0("../outputs/age.", "Beck2014" ,".Rda"))

## Epochs (rarefied to 3 taxa only)
out2 <- run.all.disparity.wrapper(morphospace, tree, type = "Epoch", FADLAD, inc.nodes = TRUE,
                        bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var",
                        rarefaction = FALSE)

## Save results to output folder
save(out2, file = paste0("../outputs/epoch.", slug ,".Rda"))
```