### 2017-11-28 analyses/tmp-time-slice-analyses.Rmd
renamed tmp for pre-merge

---
title: "Time slice analyses"
author: "Natalie Cooper and Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 6
    fig_height: 6
---

This code runs the analyses in the paper. Below we lay out the process in detail for one dataset, and then provide the code for the other datasets below this. Note that much of this uses functions in the `simulations.R` file.

```{r}
set.seed(123)
```

Load packages and functions
-----------------------------------

Make sure you work with `dispRity` 0.4 and TG's version of `Claddis` (much faster than GL's).

```{r, eval = TRUE, message = FALSE}
# The latest version of devtools and ape
library(devtools)
library(ape)

## The latest version of Claddis (on development on TG branch)
## install_github("TGuillerme/Claddis")
library(Claddis)

## The latest version of dispRity on release branch #TG: master is under many changes!
## install_github("TGuillerme/dispRity", ref = "release")
library(dispRity)

## Load functions using source
# Doesn't always work... Rmd problem?
source("../functions/simulations_fun.R")
```

Loading the data
-----------------------------------

To run `dispRity` analyses we need the **morphospace**, the **phylogeny** (with a root age and node labels) and the **first and last occurrence dates** (FADLAD). The tree and FADLAD are in the `data/` folder. To construct the morphospace we need to use the `distance_matrix` files we created in `01-extract-data-for analyses.Rmd` and stored in `data/processed`.

```{r}
## Choose a slug
slug <- "Beck2014"

## Loading the discrete morphological matrix
## This is just to match the names with the tree
matrix <- ReadMorphNexus(paste0("../data/matrices/", slug, ".nex"))

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Loading the tree
tree <- read.nexus(paste0("../data/trees/", slug, ".tre"))
## Adding node labels and a root age (max tree age date)
tree <- makeNodeLabel(tree, method = "number", prefix = "n")
tree$root.time <- max(tree.age(tree)[, 1])

## Cleaning the tree to remove things that aren't in the matrix 
cleaned_data <- clean.data(matrix$matrix, tree)
matrix$matrix <- cleaned_data$matrix
tree <- cleaned_data$tree

## Loading the distance matrices
## See "01-extract-data-for analyses.Rmd" for how we made these
load(paste0("../data/processed/distance_matrix.", slug ,".Rda"))

## Ordinating the matrix
morphospace <- cmdscale(matrix_dist, k = nrow(matrix_dist) - 2, add = TRUE)$points

```

Scenarios
--------------------------------

We used the data to explore these different scenarios:

1. Stratigraphy. This is the traditional method, where all the species within each stratigraphic period are included in the disparity calculation. This often leads to bins of unequal duration.

2. Equally sized time bins. 
    i. Where the *duration* of the bin is equal to the median duration of the stratigraphic period.
    ii. Where the *number* of bins is equal to the number of stratigraphic periods.
    iii. Where the *number* of bins is either 2, 5, 10, 15, or 20.

3. Time slicing. 
    i. Where the *interval* between slices is equal to the median duration of the stratigraphic period.
    ii. Where the *number* of slices is equal to the number of stratigraphic periods.
    iii. Where the *number* of slices is either 2, 5, 10, 15, or 20.

Running the analyses and saving the outputs
--------------------------------------------

These analyses use a series of functions in `functions\`. These essentially wrap the existing `dispRity` functions so that running lots of models and extracting lots of data is not such a big issue.

> TG: for the following pipeline, I've reduced the number of rarefaction that highly increased the calculation time.
We now only calculate the disparity for the non-rarefied subsets and the subsets rarefied to contain only 3 taxa (the lowest level of rarefaction).
We can use the full rarefaction if we want to plot rarefaction curves later on.

```{r, eval = TRUE, warning = FALSE}
## Run the function for Beck data
slug <- "BeckLee"

## Ages (rarefied to 3 taxa only)
out1 <- run.all.disparity.wrapper(morphospace, tree, type = "Age", FADLAD, inc.nodes = TRUE,
                                bootstraps = 100, metric = c(sum, variances),
                                metric.name = "sum_var", rarefaction = FALSE)
#readr::write_csv(out1, path = paste0("../outputs/age", "_", slug, ".csv"))

## Epochs (rarefied to 3 taxa only)
out2 <- run.all.disparity.wrapper(morphospace, tree, type = "Epoch", FADLAD, inc.nodes = TRUE,
                        bootstraps = 100, metric = c(sum, variances), metric.name = "sum_var",
                        rarefaction = FALSE)
#readr::write_csv(out2, path = paste0("../outputs/epoch", "_", slug, ".csv"))
```

Note that you can use the previous output format (the data tables) as before by accessing them through `out$results$stratigraphy/duration/number`.
Writing them as a csv will require some thinking since they don't have the same number of columns (though it's still doable I guess).
Also, if you want to plot things through `ggplot`, you can use the values in these data tables!

Examining the output 
--------------------

The results as a table

```{r, eval = TRUE, echo = FALSE}
knitr::kable(out2$results$stratigraphy, caption = "Disparity per stratigraphic time: unequal time bins and non-equidistant time slices")
knitr::kable(out2$results$duration, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the median stratigraphic epoch duration")
knitr::kable(out2$results$number, caption = "Disparity per average stratigraphic time: equal time bins and equidistant time slices based on the number of stratigraphic epochs")
```

This is not really digest and we should find a better way to present that.
Anyway, these boring tables are always good for the supplementaries.

## Plotting the results `dispRity` style

Here is an example of the results using `plot.dispRity`.
The upper panel is with the CIs and the lower one is the same without the CIs.

### The results for the Age type

```{r, eval = TRUE, fig.width = 12, fig.height = 18, echo = TRUE}
op <- par(bty = "n", mfrow = c(3,2))
## Principal colors
colors <- palette()[1:7][-1] #TG: that can be made fancier
colors[7] <- "black"
## Color combination (for the CI)
colors_CI <- c("pink", "lightgreen", "lightblue", "lightcyan", "lightpink", "beige", "lightgray")
## y limits
ylim <- range(unlist(lapply(out1$results, function(X) return(c(min(X[,5], na.rm = TRUE), max(X[,5],
  na.rm = TRUE))))))
## Zooming
ylim_zoom <- c(20, 35)

## Plotting the different methods (with/without CIs)
plot.results.dispRity(out1$object, method = 1, colors = colors, colors_CI = colors_CI,
  main = "Stratigraphy", ylim = ylim, CI = TRUE, legend = TRUE)
plot.results.dispRity(out1$object, method = 1, colors = colors, colors_CI = colors_CI,
  main = "Stratigraphy (no CI)", ylim = ylim_zoom, CI = FALSE, time.bins.line = TRUE)

plot.results.dispRity(out1$object, method = 2, colors = colors, colors_CI = colors_CI,
  main = "Duration", ylim = ylim, CI = TRUE)
plot.results.dispRity(out1$object, method = 2, colors = colors, colors_CI = colors_CI,
  main = "Duration (no CI)", ylim = ylim_zoom, CI = FALSE, time.bins.line = TRUE)

plot.results.dispRity(out1$object, method = 3, colors = colors, colors_CI = colors_CI,
  main = "Number", ylim = ylim, CI = TRUE)
plot.results.dispRity(out1$object, method = 3, colors = colors, colors_CI = colors_CI,
  main = "Number (no CI)", ylim = ylim_zoom, CI = FALSE, time.bins.line = TRUE)
```

### The results for the Epoch type

```{r, eval = TRUE, fig.width = 12, fig.height = 18, echo = TRUE}
op <- par(bty = "n", mfrow = c(3,2))
## Principal colors
colors <- palette()[1:7][-1] #TG: that can be made fancier
colors[7] <- "black"
## Color combination (for the CI)
colors_CI <- c("pink", "lightgreen", "lightblue", "lightcyan", "lightpink", "beige", "lightgray")
## y limits
ylim <- range(unlist(lapply(out2$results, function(X) return(c(min(X[,5], na.rm = TRUE), max(X[,5],
  na.rm = TRUE))))))

## Zooming
ylim_zoom <- c(20, 35)

## Plotting the different methods (with/without CIS)
plot.results.dispRity(out2$object, method = 1, colors = colors, colors_CI = colors_CI,
  main = "Stratigraphy", ylim = ylim, CI = TRUE, legend = TRUE)
plot.results.dispRity(out2$object, method = 1, colors = colors, colors_CI = colors_CI,
  main = "Stratigraphy (no CI)", ylim = ylim_zoom, CI = FALSE, time.bins.line = TRUE)

plot.results.dispRity(out2$object, method = 2, colors = colors, colors_CI = colors_CI,
  main = "Duration", ylim = ylim, CI = TRUE)
plot.results.dispRity(out2$object, method = 2, colors = colors, colors_CI = colors_CI,
  main = "Duration (no CI)", ylim = ylim_zoom, CI = FALSE, time.bins.line = TRUE)

plot.results.dispRity(out2$object, method = 3, colors = colors, colors_CI = colors_CI,
  main = "Number", ylim = ylim, CI = TRUE)
plot.results.dispRity(out2$object, method = 3, colors = colors, colors_CI = colors_CI,
  main = "Number(no CI)", ylim = ylim_zoom, CI = FALSE, time.bins.line = TRUE)

```

## Plotting the results histogram style

Another way for the plots would be kind of our histogram representation (you can certainly use `ggplot` here, below is just an example):

### The results per method

```{r, fig.width = 12, fig.height = 18}
## Set the list of available models
models <- list("equalbins", "acctran", "deltran", "random", "proximity", "punctuated", "gradual")

## Get the stratigraphic, duration and number data for the ages
strat_data_ages <- lapply(models, make.class.histogram, out1$results, method = 1)
durat_data_ages <- lapply(models, make.class.histogram, out1$results, method = 2)
numbe_data_ages <- lapply(models, make.class.histogram, out1$results, method = 3)

## Get the stratigraphic, duration and number data for the epochs
strat_data_epoc <- lapply(models, make.class.histogram, out2$results, method = 1)
durat_data_epoc <- lapply(models, make.class.histogram, out2$results, method = 2)
numbe_data_epoc <- lapply(models, make.class.histogram, out2$results, method = 3)

## The lower age and the lower disparity limit are arbitrarily fixed to 0 and 15 respectively.
lower_age <- 0; lower_lim <- 15
#TG: might want to change that for other datasets!

## limits functions
get.break <- function(x) return(x$breaks)
get.count <- function(x) return(x$counts)
## Ages limits
xlim_ages <- c(max(unlist(c(sapply(strat_data_ages, get.break), sapply(durat_data_ages, get.break),
  sapply(numbe_data_ages, get.break)))), lower_age)
ylim_ages <- c(lower_lim, max(unlist(c(sapply(strat_data_ages, get.count),
  sapply(durat_data_ages, get.count), sapply(numbe_data_ages, get.count)))))
## Epoch limits
xlim_epoc <- c(max(unlist(c(sapply(strat_data_epoc, get.break), sapply(durat_data_epoc, get.break),
  sapply(numbe_data_epoc, get.break)))), lower_age)
ylim_epoc <- c(lower_lim, max(unlist(c(sapply(strat_data_epoc, get.count),
  sapply(durat_data_epoc, get.count), sapply(numbe_data_epoc, get.count)))))

## Colors
colors <- palette()[1:7]
colors[1] <- "grey"

## Graphics
op <- par(bty = "n", mfrow = c(3,2))

## Plot the stratigraphy - ages
plot(strat_data_ages[[1]], xlim = xlim_ages, ylim = ylim_ages , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Stratigraphy - ages")
## Plot the slicing
for(slice in 2:7) {
  lines(strat_data_ages[[slice]]$breaks, strat_data_ages[[slice]]$counts, col = colors[slice])
}

## Adding the legend
legend(x = "bottomleft", legend = c("time bins", "acctran", "deltran", "random", "proximity"),
  col = colors[1:7], lty = 1, bg = "white")

## Plot the stratigraphy - epoch
plot(strat_data_epoc[[1]], xlim = xlim_epoc, ylim = ylim_epoc , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Stratigraphy - Epoch")
## Plot the slicing
for(slice in 2:7) {
  lines(strat_data_epoc[[slice]]$breaks, strat_data_epoc[[slice]]$counts, col = colors[slice])
}

## Plot the duration - ages
plot(durat_data_ages[[1]], xlim = xlim_ages, ylim = ylim_ages , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Duration - ages")
## Plot the slicing
for(slice in 2:7) {
  lines(durat_data_ages[[slice]]$breaks, durat_data_ages[[slice]]$counts, col = colors[slice])
}

## Plot the duration - epoch
plot(durat_data_epoc[[1]], xlim = xlim_epoc, ylim = ylim_epoc , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Duration - Epoch")
## Plot the slicing
for(slice in 2:7) {
  lines(durat_data_epoc[[slice]]$breaks, durat_data_epoc[[slice]]$counts, col = colors[slice])
}

## Plot the number - ages
plot(numbe_data_ages[[1]], xlim = xlim_ages, ylim = ylim_ages , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Number - ages")
## Plot the slicing
for(slice in 2:7) {
  lines(numbe_data_ages[[slice]]$breaks, numbe_data_ages[[slice]]$counts, col = colors[slice])
}

## Plot the number - epoch
plot(numbe_data_epoc[[1]], xlim = xlim_epoc, ylim = ylim_epoc , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Number - Epoch")
## Plot the slicing
for(slice in 2:7) {
  lines(numbe_data_epoc[[slice]]$breaks, numbe_data_epoc[[slice]]$counts, col = colors[slice])
}
```

Some of the plots are a bit off set and everything but it gives us an idea

### The results per model

Something like that but with transparency:

```{r, fig.width = 6, fig.height = 6}
plot(strat_data_epoc[[1]], xlim = xlim_epoc, ylim = ylim_epoc , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[1], main = "Epoch")
plot(durat_data_epoc[[1]], xlim = xlim_epoc, ylim = ylim_epoc , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[2], main = "", add = TRUE)
plot(numbe_data_epoc[[1]], xlim = xlim_epoc, ylim = ylim_epoc , xlab = "Time (Mya)",
  ylab = "Sum of variance", col = colors[3], main = "", add = TRUE)
```