### 2017-11-29 manuscript/tables/datasets.tex
Add draft table of dataset details

% Dataset descriptions table

\begin{tabular}[t]{l p{3cm} l l l l}

\hline
\textbf{Dataset} & \textbf{Group} & \textbf{$N_{taxa}$} & \textbf{$N_{characters}$} & \textbf{Age range} & \textbf{Reference} \\
\hline
Beck2014 & Eutherian mammals & 102 & 410 & 170 Ma - present & X \\
%-----------------------------------------------------------
Bapst2016 & theropods & 102 & 410 & 170 Ma - present & X \\
%-------------------------------------------------------
Brusatte2014 & theropods & 102 & 410 & 170 Ma - present & X \\
%-------------------------------------------------------
Heath2016 & bears & 102 & 410 & 170 Ma - present & X \\
%-------------------------------------------------------
Wright2017 & crinoids & 102 & 410 & 170 Ma - present & X \\
%-------------------------------------------------------
\hline
\end{tabular}