### 2017-12-07 analyses/04-figures.Rmd
Finished DTT figure

---
title: "Disparity through time - simplified and for appendix"
author: "Natalie Cooper"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

```{r, message = FALSE}
library(tidyverse)
library(gridExtra)
library(grid)
source("../functions/multiplot_fun.R")
```

```{r}
slug <- "Beck2014"
```

Load the data

```{r}
out1 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

strat <- out1$results$stratigraphy
duration <- out1$results$duration
number <- out1$results$number
```

Get the data into the correct format to plot

```{r}
# Stratigraphy
# Get ages for the bins, go for the midpoints
x <- str_split(strat$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

# Modify data - add midpoints, remove missing data, filter to just two models,
# make disparity relative by dividing by max values
strat2 <- 
  strat %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Duration
x <- str_split(duration$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

duration2 <- 
  duration %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Number
x <- str_split(number$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

number2 <- 
  number %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))
```

Make plots in preparation for larger grid plot below

```{r}

beck_base <- 
  ggplot(strat2, aes(y = relative.disparity, x = time, colour = model)) +
  theme_classic(base_size = 12) + 
  scale_y_continuous(breaks = seq(0, 1, 0.5), 
                     labels = seq(0, 1, 0.5)) +
  scale_x_reverse(limits = c(100, 0),
                  breaks = c(seq(100, 0, by = -50))) +
  expand_limits(x = 0, y = c(0,1)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

beck_s <- 
  beck_base +
  geom_point(data = strat2, size = 2, alpha = 0.8) +
  geom_errorbar(data = strat2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  ggtitle("Stratigraphy") +
  theme(plot.title = element_text(size = 14, hjust = 0.5)) +
  annotate("text", x = 75, y = 0.15, label = "Beck2014")

beck_d <- 
  beck_base +
  geom_point(data = duration2, size = 2, alpha = 0.8) +
  geom_errorbar(data = duration2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  ggtitle("Duration") +
  theme(plot.title = element_text(size = 14, hjust = 0.5))

beck_n <- 
  beck_base +
  geom_point(data = number2, size = 2, alpha = 0.8) +
  geom_errorbar(data = number2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  ggtitle("Number") +
  theme(plot.title = element_text(size = 14, hjust = 0.5))

```

Pick the slug

```{r}
slug <- "Brusatte2014"
```

Load the data

```{r}
out1 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

strat <- out1$results$stratigraphy
duration <- out1$results$duration
number <- out1$results$number
```

Get the data into the correct format to plot

```{r}
# Stratigraphy
# Get ages for the bins, go for the midpoints
x <- str_split(strat$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

# Modify data - add midpoints, remove missing data, filter to just two models
strat2 <- 
  strat %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Duration
x <- str_split(duration$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

duration2 <- 
  duration %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Number
x <- str_split(number$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

number2 <- 
  number %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))
```

Make plots in preparation for larger grid plot below

```{r}

brusatte_base <- 
  ggplot(strat2, aes(y = relative.disparity, x = time, colour = model)) +
  theme_classic(base_size = 12) + 
  scale_y_continuous(breaks = seq(0, 1, 0.5), 
                     labels = seq(0, 1, 0.5)) +
  scale_x_reverse(limits = c(100, 0),
                  breaks = c(seq(100, 0, by = -50))) +
  expand_limits(x = 0, y = c(0,1)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

brusatte_s <- 
  brusatte_base +
  geom_point(data = strat2, size = 2, alpha = 0.8) +
  geom_errorbar(data = strat2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  annotate("text", x = 70, y = 0.15, label = "Brusatte2014")

brusatte_d <- 
  brusatte_base +
  geom_point(data = duration2, size = 2, alpha = 0.8) +
  geom_errorbar(data = duration2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8)

brusatte_n <- 
  brusatte_base +
  geom_point(data = number2, size = 2, alpha = 0.8) +
  geom_errorbar(data = number2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8)

```

```{r}
slug <- "Bapst2016"
```

Load the data

```{r}
out1 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

strat <- out1$results$stratigraphy
duration <- out1$results$duration
number <- out1$results$number
```

Get the data into the correct format to plot

```{r}
# Stratigraphy
# Get ages for the bins, go for the midpoints
x <- str_split(strat$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

# Modify data - add midpoints, remove missing data, filter to just two models
strat2 <- 
  strat %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Duration
x <- str_split(duration$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

duration2 <- 
  duration %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Number
x <- str_split(number$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

number2 <- 
  number %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))
```

Make plots in preparation for larger grid plot below

```{r}
bapst_base <- 
  ggplot(strat2, aes(y = relative.disparity, x = time, colour = model)) +
  theme_classic(base_size = 12) + 
  scale_y_continuous(breaks = seq(0, 1, 0.5), 
                     labels = seq(0, 1, 0.5)) +
  scale_x_reverse(limits = c(100, 0),
                  breaks = c(seq(100, 0, by = -50))) +
  expand_limits(x = 0, y = c(0,1)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

bapst_s <- 
  bapst_base +
  geom_point(data = strat2, size = 2, alpha = 0.8) +
  geom_errorbar(data = strat2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  annotate("text", x = 75, y = 0.15, label = "Bapst2016")

bapst_d <- 
  bapst_base +
  geom_point(data = duration2, size = 2, alpha = 0.8) +
  geom_errorbar(data = duration2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8)

bapst_n <- 
  bapst_base +
  geom_point(data = number2, size = 2, alpha = 0.8) +
  geom_errorbar(data = number2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8)

```

```{r}
slug <- "Wright2017"
```

Load the data

```{r}
out1 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

strat <- out1$results$stratigraphy
duration <- out1$results$duration
number <- out1$results$number
```

Get the data into the correct format to plot

```{r}
# Stratigraphy
# Get ages for the bins, go for the midpoints
x <- str_split(strat$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

# Modify data - add midpoints, remove missing data, filter to just two models
strat2 <- 
  strat %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Duration
x <- str_split(duration$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

duration2 <- 
  duration %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))

# Number
x <- str_split(number$subset, " - ") 
y <- map(x, as.numeric)
z <- map(y, mean)

number2 <- 
  number %>%
  mutate(time = unlist(z)) %>%
  filter(bs.median != 0 & !is.na(bs.median)) %>%
  filter(model == "equalbins" | model == "proximity" | model == "gradual.split") %>%
  mutate(relative.disparity = bs.median/max(bs.median)) %>%
  mutate(relative.lwr = `2.5%`/max(bs.median)) %>%
  mutate(relative.upr = `97.5%`/max(bs.median))
```

Make plots in preparation for larger grid plot below

```{r}

wright_base <- 
  ggplot(strat2, aes(y = relative.disparity, x = time, colour = model)) +
  theme_classic(base_size = 12) + 
  scale_y_continuous(breaks = seq(0, 1, 0.5), 
                     labels = seq(0, 1, 0.5)) +
  scale_x_reverse(limits = c(470, 370),
                  breaks = c(seq(470, 370, by = -50))) +
  expand_limits(x = 0, y = c(0,1)) +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")

wright_s <- 
  wright_base +
  geom_point(data = strat2, size = 2, alpha = 0.8) +
  geom_errorbar(data = strat2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  annotate("text", x = 440, y = 0.15, label = "Wright2017")

wright_d <- 
  wright_base +
  geom_point(data = duration2, size = 2, alpha = 0.8) +
  geom_errorbar(data = duration2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8) +
  scale_x_reverse(limits = c(470, 370),
                  breaks = c(seq(470, 370, by = -100)))

wright_n <- 
  wright_base +
  geom_point(data = number2, size = 2, alpha = 0.8) +
  geom_errorbar(data = number2, aes(ymin = relative.lwr, ymax = relative.upr), width = 1, alpha = 0.8)

```

Finally plot them all together

```{r}
# Using modified multiplot function
# written by Scott Chamberlain

pdf("../manuscript/figures/fig-dtt.pdf")  
  multiplot(beck_s + theme(plot.margin=unit(c(5.5, 5.5, 5.5, 25),"points")),
            brusatte_s + theme(plot.margin=unit(c(5.5, 5.5, 5.5, 25),"points")), 
            bapst_s + theme(plot.margin=unit(c(5.5, 5.5, 5.5, 25),"points")),
            wright_s + theme(plot.margin=unit(c(5.5, 5.5, 12, 25),"points")), 
            beck_d, brusatte_d, bapst_d, 
            wright_d + theme(plot.margin=unit(c(5.5, 5.5, 12, 5.5),"points")),
            beck_n,
            brusatte_n,
            bapst_n,
            wright_n + theme(plot.margin=unit(c(5.5, 5.5, 12, 5.5),"points")),
            cols = 3, 
            labs = list("Time (MYA)", 
                        "Relative disparity"),
            labpos=list(c(0.5, 0.015), c(0.03, 0.5)))
dev.off()
```