### 2017-12-08 analyses/06-table-wilcox.Rmd
finish wilcox tables code

---
title: "Wilcoxon results tables - text and appendix"
author: "Natalie Cooper"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 12
---

# Compiling the tables for the paper

Loading the packages

```{r}
library(xtable)
library(Claddis)
library(ape)
library(dispRity)
library(tidyverse)
```

Loading the data:

```{r}
## Getting the data
slugs <- list("Beck2014", "Brusatte2014", "Bapst2016", "Wright2017")

## Wilcox results
wilk <- lapply(slugs, function(slug) read.csv(paste0("../outputs/wilcox.", slug, ".csv"), 
                                              header = FALSE))

```

Manipulate the data into the format we need.

```{r}
# Stick p values and W values together
# Repalce p values with *
p.values.function <- function(data){ 
  for (i in 1:length(data$V2)){
    if(data$V2[i] < 0.05 & data$V2[i] > 0.01){data$V2[i] <- "*"}
    if(data$V2[i] < 0.01 & data$V2[i] > 0.001){data$V2[i] <- "**"}
    if(data$V2[i] < 0.001){data$V2[i] <- "***"}
    if(data$V2[i] > 0.05){data$V2[i] <- ""}
  }
  data$V6 <- paste0(data$V1, data$V2)
return(data)
}

# Extract proximity and gradual split models
# Remove excess columns
# Add model column name
fix.wilcox <- function(data){
  data %>%
  select(-c(V1,V2)) %>%
  filter(V3 == "proximity" | V3 == "gradual.split") %>%
  spread(V4, V6) %>%
  arrange(V5) %>%
  rename(model = V3) %>%
  rename(period = V5)  
}

wilk2 <- purrr::map(wilk, p.values.function)
wilk3 <- purrr::map_df(wilk2, fix.wilcox)
wilk3$dataset <- c(rep(unlist(slugs)[1], 4), rep(unlist(slugs)[2], 4), 
                   rep(unlist(slugs)[3], 4), rep(unlist(slugs)[4], 4))

```

## Create table

```{r}
## variable names
colnames <- c("Dataset", "Period", "Model", "Stratigraphy", "Duration", "Number")

## Formatting the table
table_out <- data.frame(wilk3$dataset, wilk3$period, wilk3$model, wilk3$stratigraphy,
                        wilk3$duration, wilk3$number)
colnames(table_out) <- colnames

## Table caption
table_caption <- "Wilcoxon results"

sink(file = "../manuscript/tables/raw_wilcox.tex")
print(xtable(table_out, caption = table_caption), include.rownames = FALSE)
sink()
```

## Repeat for appendix tables with all models

Manipulate the data into the format we need.

```{r}
# Remove excess columns
# Add model column name
fix.wilcox.appendix <- function(data){
  data %>%
  select(-c(V1,V2)) %>%
  spread(V4, V6) %>%
  arrange(V5) %>%
  rename(model = V3) %>%
  rename(period = V5)
}

wilk2 <- purrr::map(wilk, p.values.function)
wilk3 <- purrr::map_df(wilk2, fix.wilcox.appendix)
wilk3$dataset <- c(rep(unlist(slugs)[1], 12), rep(unlist(slugs)[2], 12), 
                   rep(unlist(slugs)[3], 12), rep(unlist(slugs)[4], 12))

```

## Create table for all data

```{r}
## variable names
colnames <- c("Dataset", "Period", "Model", "Stratigraphy", "Duration", "Number")

## Formatting the table
table_out <- data.frame(wilk3$dataset, wilk3$period,
                        wilk3$model, wilk3$stratigraphy,
                        wilk3$duration, wilk3$number)
colnames(table_out) <- colnames

## Table caption
table_caption <- "Wilcoxon results for appendix"

sink(file = "../manuscript/tables/raw_wilcox_appendix.tex")
print(xtable(table_out, caption = table_caption), include.rownames = FALSE)
sink()
```