### 2017-12-08 manuscript/tables/table_wilcox_appendix.tex

% latex table generated in R 3.4.2 by xtable 1.8-2 package
% Fri Dec  8 17:06:33 2017
\begin{table}[!htbp]
\centering
\begin{tabular}{lllccc}
  \hline
\textbf{Dataset} & \textbf{Period} & \textbf{Model} & \textbf{Stratigraphy} & \textbf{Duration} & \textbf{Number} \\ 
  \hline
Beck2014 & Age & acctran & 39*** & 76 & 11 \\ 
  Beck2014 & Age & deltran & 188*** & 194*** & 171 \\ 
  Beck2014 & Age & equal.split & 91 & 119*** & 47 \\ 
  Beck2014 & Age & gradual.split & 111 & 115*** & 65*** \\ 
  Beck2014 & Age & proximity & 105 & 83 & 68*** \\ 
  Beck2014 & Age & random & 97 & 104*** & 45 \\ 
  Beck2014 & Epoch & acctran & 14 & 10 & 14 \\ 
  Beck2014 & Epoch & deltran & 21 & 45*** & 41*** \\ 
  Beck2014 & Epoch & equal.split & 21 & 40*** & 42*** \\ 
  Beck2014 & Epoch & gradual.split & 21 & 39 & 43*** \\ 
  Beck2014 & Epoch & proximity & 21 & 36 & 32 \\ 
  Beck2014 & Epoch & random & 21 & 37 & 45*** \\ 
  Brusatte2014 & Age & acctran & 27*** & 28 & 28*** \\ 
  Brusatte2014 & Age & deltran & 27*** & 29 & 31*** \\ 
  Brusatte2014 & Age & equal.split & 28*** & 58*** & 50*** \\ 
  Brusatte2014 & Age & gradual.split & 28*** & 61*** & 52*** \\ 
  Brusatte2014 & Age & proximity & 27*** & 31 & 28*** \\ 
  Brusatte2014 & Age & random & 27*** & 27 & 27*** \\ 
  Brusatte2014 & Epoch & acctran & 0 & 5*** & 5 \\ 
  Brusatte2014 & Epoch & deltran & 0 & 5*** & 5 \\ 
  Brusatte2014 & Epoch & equal.split & 3 & 6 & 6 \\ 
  Brusatte2014 & Epoch & gradual.split & 3 & 6 & 6 \\ 
  Brusatte2014 & Epoch & proximity & 0 & 5*** & 5 \\ 
  Brusatte2014 & Epoch & random & 0 & 5*** & 5 \\ 
  Bapst2016 & Age & acctran & 45*** & 47 & 72*** \\ 
  Bapst2016 & Age & deltran & 55*** & 46 & 78*** \\ 
  Bapst2016 & Age & equal.split & 93 & 147*** & 153 \\ 
  Bapst2016 & Age & gradual.split & 93 & 153 & 165 \\ 
  Bapst2016 & Age & proximity & 57*** & 47 & 75*** \\ 
  Bapst2016 & Age & random & 57*** & 48 & 81*** \\ 
  Bapst2016 & Epoch & acctran & 2 & 0*** & 8 \\ 
  Bapst2016 & Epoch & deltran & 2 & 0*** & 9 \\ 
  Bapst2016 & Epoch & equal.split & 4 & 6 & 13 \\ 
  Bapst2016 & Epoch & gradual.split & 4 & 6 & 12 \\ 
  Bapst2016 & Epoch & proximity & 2 & 0*** & 8 \\ 
  Bapst2016 & Epoch & random & 2 & 1*** & 8 \\ 
  Wright2017 & Age & acctran & 146*** & 146 & 84 \\ 
  Wright2017 & Age & deltran & 162*** & 138 & 101 \\ 
  Wright2017 & Age & equal.split & 151*** & 160 & 105 \\ 
  Wright2017 & Age & gradual.split & 152*** & 155 & 116 \\ 
  Wright2017 & Age & proximity & 160*** & 175*** & 101 \\ 
  Wright2017 & Age & random & 150*** & 147 & 111 \\ 
  Wright2017 & Epoch & acctran & 25 & 20 & 18 \\ 
  Wright2017 & Epoch & deltran & 27 & 26 & 25 \\ 
  Wright2017 & Epoch & equal.split & 29 & 30 & 25 \\ 
  Wright2017 & Epoch & gradual.split & 28 & 29 & 21 \\ 
  Wright2017 & Epoch & proximity & 23 & 28 & 18 \\ 
  Wright2017 & Epoch & random & 28 & 23 & 17 \\ 
   \hline
\end{tabular}
\caption{Wilcoxon results for appendix} 
\end{table}