### 2017-12-15 analyses/09-appendix-trees-figures.Rmd
Add code to make strap trees

---
title: "Trees for the appendix"
author: "Natalie Cooper"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Load the packages 
```{r}
library(strap)
```

Load the data

```{r}
## Choose a slug
slug <- "Beck2014"
#slug <- "Brusatte2014"
#slug <- "Bapst2016"
#slug <- "Wright2017"

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Loading the tree
tree <- read.nexus(paste0("../data/processed/tree.", slug, ".tre"))
## Add root age (max tree age date)
## Remember that for Wright this is 485.4
tree$root.time <- max(tree.age(tree)[, 1])
```

Plot the trees and save as PDF
```{r}
pdf(paste0("../manuscript/figures/fig-tree-", slug, "-appendix.pdf"), width = 12, height = 12)  
geoscalePhylo(tree = tree, ages = FADLAD)
dev.off()
```