### 2017-12-18 analyses/09-appendix-trees-figures.Rmd
Strap plots

---
title: "Trees for the appendix"
author: "Natalie Cooper"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Load the packages

```{r}
library(strap)
library(dispRity)
library(ape)
```

Load the data

```{r}
## Choose a slug
slug <- "Beck2014"
# slug <- "Brusatte2014"
# slug <- "Bapst2016" #ok
# slug <- "Wright2017" #ok

## Loading the tree
tree <- read.nexus(paste0("../data/processed/tree.", slug, ".tre"))

## Loading the FADLAD
FADLAD <- read.csv(paste0("../data/FADLAD/", slug, ".csv"), row.names = 1, header = TRUE)

## Remove the taxa not present in the tree but in the FADLAD
absent_in_tree <- clean.data(FADLAD, tree)$dropped_rows
if(!any(is.na(absent_in_tree))) {
    FADLAD <- FADLAD[-match(absent_in_tree, rownames(FADLAD)),]
}

## Getting the tree ages
tree_ages <- tree.age(tree)

## Scaling with the latest FADLAD
tree$root.time <- max(tree_ages[,1] + min(FADLAD[,2]))

## Correct the root age for Wright2017
tree$root.time <- ifelse(slug == "Wright2017", 485.4, tree$root.time)

## Updating the tree age
tree_ages <- tree.age(tree)

## Updating the FADLAD (if missing tips)
if(nrow(FADLAD) !=  Ntip(tree)) {
    ## Selecting the missing taxa
    missing_taxa <- which(is.na(match(tree_ages$elements[1:Ntip(tree)], rownames(FADLAD))))
    ## Creating a FADLAD table for the missing taxa (FAD = LAD = tip age)
    FADLAD_tmp <- data.frame(FAD = tree_ages$ages[missing_taxa], LAD = tree_ages$ages[missing_taxa],
                             row.names = tree_ages$elements[missing_taxa])
    ## Combining both.data frames
    FADLAD <- rbind(FADLAD, FADLAD_tmp)
}
```

Plot the trees and save as PDF

```{r}
pdf(paste0("../manuscript/figures/fig-tree-", slug, "-appendix.pdf"), width = 12, height = 12) 
## Adjust plot sizes (for offset labels with e.g.: x.lim = c(350, 485.4) and cex.tip = 0.5))
geoscalePhylo(tree = tree, ages = FADLAD)
dev.off()
```