### 2017-12-18 analyses/12-appendix-disparity-tables.Rmd
added table compiler (96 pages of appendix!)

---
title: "Disparity tables for the appendix q"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

# Building the disparity tables for each dataset

```{r}
library(dispRity)
library(xtable)

## Function for only selecting the first occurrence of each level (for fancy tables)
level.selector <- function(column, data) {
    ## Selecting the data
    data_col <- data[, column]

    ## Selecting the levels
    all_levels <- levels(as.factor(data_col))

    ## Replacing them in the column
    for(one_level in 1:length(all_levels)) {
        ## Selecting the levels occurrences
        level_occurrences <- data_col %in% all_levels[one_level]
        ## Selecting the first of each
        first_level_occurrences <- sapply(seq(2:length(level_occurrences[-1])), function(X, test)
            return(ifelse((level_occurrences[X] == TRUE && level_occurrences[X-1] == FALSE)
                , TRUE, FALSE)), level_occurrences)

        ## Replacing the values in data_col
        data_col[which(first_level_occurrences)] <- paste0(data_col[which(first_level_occurrences)],
                                                           "@1st")
    }

    ## Remove the non first entries
    data_col[-grep("@1st", data_col)] <- ""
    data_col <- gsub("@1st", "", data_col)

    return(data_col)
}

## Rounding slices/bins values
round.slices <- function(slice, digits = 3) {
    options(warn = -1)
    ## Convert into numeric
    slice_value <- as.numeric(slice)
    options(warn = 0)

    if(is.na(slice_value)) {
        ## Is the value simply NA or a bin
        if(grep("-", slice) > 0) {
            ## Separating both values
            bin_values <- unlist(strsplit(slice, split = " - "))
            ## Rounding the values
            bin_values <- round(as.numeric(bin_values), digits = digits)
            ## Recombining them
            return(paste(bin_values, collapse = " - "))

        } else {
            ## Slice value is a real NA
            return(as.character(NA))
        }
    } else {
        ## Slice is a slice
        return(as.character(round(as.numeric(slice_value), digits = digits)))
    }
}

```

## Generate the raw results tables for each datasets

```{r, warn = FALSE}
slugs <- c("Beck2014", "Brusatte2014", "Bapst2016", "Wright2017")

for(one_slug in 1:length(slugs)) {

    slug <- slugs[one_slug]

    ## Loading the data
    out1 <- readRDS(file = paste0("../outputs/age.", slug, ".Rda"))
    out2 <- readRDS(file = paste0("../outputs/epoch.", slug, ".Rda"))

    ## bind all the data together
    results <- rbind(do.call(rbind, out1$results), do.call(rbind, out2$results))

    ## Re-order the columns
    results <- results[,c(10,11,9, 1:8)]
    results[,1:3] <- apply(results[,1:3], 2, as.character)

    ## Empty repetition cells (for aesthetics)
    ## Stratigraphy
    results[, 1] <- level.selector(1, results)
    ## Bin_type
    results[, 2] <- level.selector(2, results)
    ## Model
    results[, 3] <- level.selector(3, results)

    ## Correcting the first slice
    results[1, c(1,2,3)] <- c("Ages", "stratigraphy", "equalbins")

    ## Rounding the slices
    results[, 4] <- sapply(results[, 4], round.slices, digits = 2)

    ## Numbering the columns
    results[, 5:11] <- apply(results[, 5:11], 2, as.numeric)

    ## Table caption
    table_caption <- paste("Disparity measurement (sum of variance) for", slug, "dataset")

    ## Save the table (as longtable)
    write(print(xtable(results, caption = table_caption, digits = 2),
                tabular.environment = "longtable", include.rownames = FALSE, floating = FALSE),
          file = paste0("../manuscript/tables/disparity_measurement-", slug, ".tex"))
}
```