### 2017-12-20 manuscript/tables/raw_datasets.tex
Edit to caption and extinctions

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Wed Dec 20 13:12:18 2017
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & Beck2014 & Brusatte2014 & Bapst2016 & Wright2017 \\ 
  \hline
Group & mammals & theropods & theropods & crinoids \\ 
  \# taxa & 106 & 152 &  89 &  42 \\ 
  \# characters & 421 & 853 & 374 &  87 \\ 
  Age range (MYA) & 171.786 - 0 & 102.535 - 0 & 141.216 - 0 & 485.4 - 372.2 \\ 
  Mass extinction (MYA) & 66 (K-Pg) & NA & NA & 443 (O-S) \\ 
  Reference &  &  &  &  \\ 
  Data reference &  &  &  &  \\ 
   \hline
\end{tabular}
\caption{Details of the datasets used in this study. Note that although Brusatte2014 and Bapst2016 cross the Cretaceaous Palaeogene (K-Pg) mass extinction there are not enough taxa post K-Pg to investigate the effects of the mass extinction on disparity.} 
\end{table}