### 2018-02-02 manuscript/WithinOrdinationACE.Rmd
pre/post ordination ACE

---
title: "Within ordination ACE"
author: "Thomas Guillerme"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 12
    fig_height: 6
---

Let's consider a square of size one in four dimensions

```{r, fig.width = 6, fig.height = 6}
## square coordinates
square <- matrix(ncol = 4, nrow = 4, byrow = TRUE,
                  data = c(c(0,0,0,0), c(0,1,0,0), c(1,1,0,0), c(1,0,0,0)))
rownames(square) <- letters[1:4]
## The 2D square
par(bty = "n")
plot(square[,c(1:2)], xlab = "Dimension 1", ylab = "Dimension 2")
```

We can then ordinate this segment using an MDS

```{r, fig.width = 6, fig.height = 6}
## Ordination (reducing only one dimension)
ord_square <- cmdscale(dist(square), k = 3)

## Plotting the two first dimensions
par(bty = "n")
plot(ord_square[,c(1:2)], xlab = "Ordination 1", ylab = "Ordination 2")
```

We can then estimate the values on the edges of the square (the mean between the pairs of summits) either within the ordination (average ordination scores) or directly from the non-ordinated square.

## Adding data within the ordination

We will now add four points that will be on the centre of each edges of the ordinated square

```{r, fig.width = 6, fig.height = 6}
## Getting the average in 3 dimensions for each edges
ab <- apply(ord_square[1:2,], 2, mean) 
bc <- apply(ord_square[2:3,], 2, mean) 
cd <- apply(ord_square[3:4,], 2, mean) 
da <- apply(ord_square[c(4,1),], 2, mean) 
ord_square_edge <- rbind(ord_square, ab, bc, cd, da)

## Plotting the two first dimensions
par(bty = "n")
plot(ord_square_edge[,c(1:2)], xlab = "Ordination 1", ylab = "Ordination 2")
```

## Adding data before the ordination

Alternatively, we can add the data in a post or pre ordination way.

```{r, fig.width = 12, fig.height = 6}
par(bty = "n", mfrow = c(1,2))

## Adding the average for each square edge
ab <- apply(square[1:2,], 2, mean) 
bc <- apply(square[2:3,], 2, mean) 
cd <- apply(square[3:4,], 2, mean) 
da <- apply(square[c(4,1),], 2, mean) 
square_edge <- rbind(square, ab, bc, cd, da)

## The 2D square (with edges center)
plot(square_edge[,c(1:2)], xlab = "Dimension 1", ylab = "Dimension 2", main = "post ordination addition")

## Ordination (reducing only one dimension)
ord_square_edge2<- cmdscale(dist(square_edge), k = 3)

## Plotting the two first dimensions
plot(ord_square_edge2[,c(1:2)], xlab = "Ordination 1", ylab = "Ordination 2", main = "pre ordination addition")
```

In the case of simple structures, addition of elements pre or post ordination has no effect on the results.
However, for more complex structures (i.e. phylogenies), this has a bigger impact on the resulting ordination:

# A phylogenetic example

Here we'll simulate a four taxa ladder tree with equal branch length (to make node estimation post-ordination comparable to pre-ordination - i.e. no effect of branch length).
We generate a realistic multistate discrete matrix (`dispRity::sim.morpho`) and estimate ancestral states for each node (`Claddis::AncStateEstMatrix`).
We also calculate the variance per axis (expressed in a probability).

```{r}
library(Claddis)
library(dispRity)
library(ape)
set.seed(1)

## Simulating a 4 taxa tree
tree <- read.tree(text = "(((t1,t2),t3),t4);")
tree$edge.length <- rep(1, 6)
## Setting up the parameters
my_rates = c(rgamma, rate = 10, shape = 5)
## Mk matrix (4*14)
matrixMk <- sim.morpho(tree, characters = 14, states = c(0.8, 0.2), model = "ER", rates = my_rates, invariant = FALSE)

## Estimating the ancestral characters
ancestors <- AncStateEstMatrix(MakeMorphMatrix(matrixMk), tree)

## Ordinating the matrix (without the ancestors)
ordination <- cmdscale(dist(matrixMk), k = nrow(matrixMk)-2)
ord_var <- apply(ordination, 2, var)
cum_var <- ord_var/sum(ord_var)
```

We can then plot both the tree and the position of each tip in the ordinated space as follows:

```{r, fig.width = 12, fig.height = 6}
## Plotting the two first axis
par(bty = "n", mfrow = c(1,2))
plot(ordination, xlab = paste0("O1 (", round(cum_var[1], 3),")"),
                 ylab = paste0("O2 (", round(cum_var[2], 3),")"),
                 xlim = c(-2.2, 2.2), ylim = c(-2.2, 2.2))

## Adding the tips names
for(tip in 1:Ntip(tree)) {
    text(ordination[tip, 1], ordination[tip, 2], labels = rownames(ordination)[tip], adj = -1)
}

## Adding the tree
plot(tree)
nodelabels(c("root", "n3", "n1.2"))
```

Note here the values of the ordination on each axis sums up to 1 since the ordination creates only two dimensions (using the $k = n-2$ rule for generating the number of axis where $k$ are the number of dimensions and $n$ the number of elements in the matrix).

### Pre/post ordination nodes

We can then add the estimates of our nodes pre or post ordination.

 * The post ordination nodes coordinates are the average values of its descendants (since branch length is always equal to one).
 * The pre ordination nodes coordinates are obtained by adding the ancestral states estimates to the input discrete morphological matrix (`matrixMk`) and re-ordinating the resulting matrix.

```{r}
## Adding the nodes to the ordination (as averages values)
n1.2 <- apply(ordination[1:2,], 2, mean) 
n3 <- apply(rbind(n1.2, ordination[3,]), 2, mean) 
root <- apply(rbind(n3, ordination[4,]), 2, mean) 
post_ordination <- rbind(ordination, n1.2, n3, root)
## Recalculating variance
post_ord_var <- apply(post_ordination, 2, var)
cum_post_ord_var <- post_ord_var/sum(post_ord_var)

## Rerunning the ordination with the ancestors
pre_ordination <- cmdscale(dist(rbind(matrixMk, ancestors)), k = nrow(rbind(matrixMk, ancestors))-2)
rownames(pre_ordination)[5:7] <- c("n1.2", "n3", "root")
## Recalculating variance
pre_ord_var <- apply(pre_ordination, 2, var)
cum_pre_ord_var <- pre_ord_var/sum(pre_ord_var)
```

For both, we recalculated the the variance per axis.
Note that due to the difference in input matrix size, the `post_ordination` matrix has still the same number of dimensions as previously (2) but the `pre_ordination` matrix has now five dimensions (7-2).

```{r, fig.width = 12, fig.height = 6}
## Plotting the two spaces
par(bty = "n", mfrow = c(1,2))

## The ordination within one
plot(post_ordination[1:4,], xlab = paste0("O1 (", round(cum_post_ord_var[1], 3),")"),
                        ylab = paste0("O2 (", round(cum_post_ord_var[2], 3),")"),
                        xlim = c(-2.2, 2.2), ylim = c(-2.2, 2.2),
                        main = "nodes added post ordination")

## Adding the tips names
for(tip in 1:(Ntip(tree))) {
    text(post_ordination[tip, 1], post_ordination[tip, 2],
        labels = rownames(post_ordination)[tip], adj = -1)
}

## The ordination without one
plot(pre_ordination[1:4,], xlab = paste0("O1 (", round(cum_pre_ord_var[1], 3),")"),
                        ylab = paste0("O2 (", round(cum_pre_ord_var[2], 3),")"),
                        xlim = c(-2.2, 2.2), ylim = c(-2.2, 2.2),
                        main = "nodes added pre ordination")

## Adding the tips names
for(tip in 1:(Ntip(tree))) {
    text(pre_ordination[tip, 1], pre_ordination[tip, 2],
        labels = rownames(pre_ordination)[tip], adj = -1)
}
```

In this case, regardless of the nodes position and/or estimation, there is a big variation in the tips position in the ordination when adding the nodes post or pre-ordination.
Note also the change in the axis variance between the post-ordination plot and the previous one (without node reconstructions).

#### Nodes position

```{r, fig.width = 12, fig.height = 6}
## Plotting the two spaces
par(bty = "n", mfrow = c(1,2))

## The ordination within one
plot(post_ordination, xlab = paste0("O1 (", round(cum_post_ord_var[1], 3),")"),
                        ylab = paste0("O2 (", round(cum_post_ord_var[2], 3),")"),
                        xlim = c(-2.2, 2.2), ylim = c(-2.2, 2.2),
                        main = "nodes added post ordination")

## Adding the tips names
for(tip in 1:(Ntip(tree))) {
    text(post_ordination[tip, 1], post_ordination[tip, 2],
        labels = rownames(post_ordination)[tip], adj = -1)
}

## The ordination without one
plot(pre_ordination, xlab = paste0("O1 (", round(cum_pre_ord_var[1], 3),")"),
                        ylab = paste0("O2 (", round(cum_pre_ord_var[2], 3),")"),
                        xlim = c(-2.2, 2.2), ylim = c(-2.2, 2.2),
                        main = "nodes added pre ordination")

## Adding the tips names
for(tip in 1:(Ntip(tree))) {
    text(pre_ordination[tip, 1], pre_ordination[tip, 2],
        labels = rownames(pre_ordination)[tip], adj = -1)
}
```